Masterclass Assignment: Real-time Global Illumination

### 2018-01-13 docs/paper.tex
finished writing paper

The final step was to implement cone tracing in the deferred shading pass. This pass calculates the indirect light contribution, specular lobe contribution, and also soft shadows.

\begin{figure}[htbp]\centering
 \includegraphics[width=1.0\linewidth]{images/BRDF}
 \caption{A series of diagrams showing how the combination of a specular cone and multiple diffuse cones can be used to approximate a material BRDF. \protect\cite{Villegas2016} The diffuse cones form a normal-oriented hemisphere to approximate the diffuse reflections, whilst the specular cone follows the reflection vector and it's aperture varies dependent on the material roughness.}
\end{figure}

@@

\begin{figure}[htbp]\centering
\begin{center}
 \begin{tabular}{|c|c|c|c|}
 \hline
 Voxel Resolution & $256^3$ & $512^3$ & $768^3$ \\
 \hline
 \hline
 Frames per second & 60 & 60 & 60 \\
 \hline
 Memory usage & 1063MiB & 2615MiB & 6628MiB \\
 \hline
 Voxelization & 2ms & 2ms & 2ms \\
 \hline
 Light Injection & 65ms & 378.5ms & 12294.5ms\\
 \hline
 Shading & 4.5ms & 4.5ms & 4.5ms \\
 \hline
 \end{tabular}
 \caption{\label{figure:performance} Averaged performance results when running with a NVIDIA GTX 1080 8GB at 2560x1080.}
\end{center}
\end{figure}

\begin{figure}[htbp]\centering
 \includegraphics[width=1.0\linewidth]{images/full_render_01}
 \includegraphics[width=1.0\linewidth]{images/full_render_02}
 \includegraphics[width=1.0\linewidth]{images/full_render_03}
 \caption{\label{figure:fullrender}Three screenshots showing the full scene with a voxel resolution of $512^3$.}
\end{figure}

\begin{figure}[htbp]\centering
 \includegraphics[width=1.0\linewidth]{images/direct_only}
 \caption{A screenshot showing just the direct lighting pass.}
\end{figure}

\begin{figure}[htbp]\centering
 \includegraphics[width=1.0\linewidth]{images/indirect_only}
 \caption{A screenshot showing just the indirect lighting pass.}
\end{figure}

\begin{figure}[htbp]\centering
 \includegraphics[width=1.0\linewidth]{images/reflection_only}
 \caption{A screenshot showing just the reflection pass.}
\end{figure}

\begin{figure}[htbp]\centering
 \includegraphics[width=1.0\linewidth]{images/reflection_only_no_roughness}
 \caption{A screenshot showing just the reflection pass, but with the minimum cone angle.}
\end{figure}

@@

\begin{figure}[htbp]\centering
 \includegraphics[width=1.0\linewidth]{images/banding.png}
 \caption{A screenshot showing banding artefacts in the reflection pass.}
\end{figure}

\begin{figure}[htbp]\centering
 \includegraphics[width=1.0\linewidth]{images/indirect_noise.png}
 \caption{A screenshot showing noise in the indirect lighting.}
\end{figure}

\begin{figure}[htbp]\centering
 \includegraphics[width=1.0\linewidth]{images/voxelization_error.png}
 \caption{A screenshot showing voxelization artefacts.}
\end{figure}

>>>>

The final step was to implement cone tracing in the deferred shading pass. This pass calculates the indirect light contribution, specular lobe contribution, and also soft shadows, using the emissive 3D texture and the textures generated by the G-Buffer.

\begin{figure}[htbp]\centering
\includegraphics[width=1.0\linewidth]{images/BRDF}
\caption{\label{figure:BRDF} A series of diagrams showing how the combination of a specular cone and multiple diffuse cones can be used to approximate a material BRDF. \protect\cite{Villegas2016} The diffuse cones form a normal-oriented hemisphere to approximate the diffuse reflections, whilst the specular cone follows the reflection vector and it's aperture varies dependent on the material roughness.}
\end{figure}

For the indirect and specular light contributions I referred to VCTRenderer \cite{Villegas2016} for example code on implementing an efficient cone tracing function, and how it can be used to approximate a BRDF as shown in Figure~\ref{figure:BRDF}. I then adapted the specular component to work with the PBR materials to ensure the specular cone aperture is physically plausible for any given viewing angle.

To calculate the the direct lighting contribution I referred to tutorial materian from LearnOpenGL \cite{learnPBR} to ensure that the PBR textures passed through the G-Buffer are combined in a physically plausible way. I referred again to VCTRenderer \cite{Villegas2016} to implement soft shadows into the direct lighting component, which involves calculating a fragment's occlusion by tracing a cone back towards the light source and accumulating the opacity values at each step.

Finally I added additional controls to the interface and passed them as uniforms to the deferred shading pass, allowing the user to control the following parameters:
\begin{itemize}
	\item The overall contributions of each lighting component.
	\item Whether to enable or disable the calculation of each lighting component, as shown in Figures~\ref{figure:direct},~\ref{figure:indirect},~\ref{figure:specular}.
	\item The cone aperture for tracing the soft shadows. Larger values can be used to estimate larger light sources.
	\item The light falloff exponent $k$ relative to the $distance$ to the light. Light falloff is calculated as $1 / (distance^k)$, so the default value of 2.0 is physically plausible.
	\item A multiplier for the specular cone aperture. Smaller values can make the scene highly reflective as shown in Figure~\ref{figure:reflection}.
\end{itemize}

@@

\begin{figure}[htbp]\centering
\begin{center}
\begin{tabular}{|c|c|c|c|}
\hline
Voxel Resolution & $256^3$ & $512^3$ & $768^3$ \\
\hline
\hline
Frames per second & 60 & 60 & 60 \\
\hline
Memory usage & 1063MiB & 2615MiB & 6628MiB \\
\hline
Voxelization & 2ms & 2ms & 2ms \\
\hline
Light Injection & 65ms & 378.5ms & 12294.5ms\\
\hline
Shading & 4.5ms & 4.5ms & 4.5ms \\
\hline
\end{tabular}
\caption{\label{figure:performance} Averaged performance results when running with a NVIDIA GTX 1080 8GB at 2560x1080.}
\end{center}
\end{figure}
\begin{figure}[htbp]\centering
\includegraphics[width=1.0\linewidth]{images/full_render_01}
\includegraphics[width=1.0\linewidth]{images/full_render_02}
\includegraphics[width=1.0\linewidth]{images/full_render_03}
\caption{\label{figure:fullrender}Three screenshots showing the full scene with a voxel resolution of $512^3$.}
\end{figure}
\begin{figure}[htbp]\centering
\includegraphics[width=1.0\linewidth]{images/direct_only}
\caption{\label{figure:direct}A screenshot showing just the direct lighting component.}
\end{figure}
\begin{figure}[htbp]\centering
\includegraphics[width=1.0\linewidth]{images/indirect_only}
\caption{\label{figure:indirect}A screenshot showing just the indirect lighting component.}
\end{figure}
\begin{figure}[htbp]\centering
\includegraphics[width=1.0\linewidth]{images/reflection_only}
\caption{\label{figure:specular}A screenshot showing just the specular component.}
\end{figure}
\begin{figure}[htbp]\centering
\includegraphics[width=1.0\linewidth]{images/reflection_only_no_roughness}
\caption{\label{figure:reflection}A screenshot showing just the specular component, but with the minimum cone aperture.}
\end{figure}

@@

There are also a number of visual artefacts, the most noticable of which being the cracks in the voxelization as seen in Figure~\ref{figure:cracks}. I understand this to be due to incorrect conservative rasterization, however I have the NVIDIA \textit{GL CONSERVATIVE RASTERIZATION NV} extension enabled and have confirmed it to be working, so I believe it could be an issue with using the extension whilst having multisampling enabled (as it is currently). I found that using a geometry shader to achieve the same result led to quite unusable artefacting.

Another artefact is quite visible noise in the indirect lighting component, as seen in Figure~\ref{figure:noise}. I am not sure what causes this, though I suspect it could be because the hemisphere of indirect cones are aligned with the normal-mapped surface normal as opposed to the surface normal recieved from the geometry, which results in excessive variance between indirect lighting calculations for neighboring fragments. Banding artefacts are visible in the specular component, as seen in Figure~\ref{figure:banding}, which occurs at certain viewing angles and at certain values for the specular aperture multiplier. I suspect this is a combination of excessive variance in surface normals, as mentioned previously, but could also be a result of limited numerical precision in calculating the specular component.

\begin{figure}[htbp]\centering
\includegraphics[width=1.0\linewidth]{images/voxelization_error.png}
\caption{\label{figure:cracks}A screenshot showing voxelization artefacts.}
\end{figure}
\begin{figure}[htbp]\centering
\includegraphics[width=1.0\linewidth]{images/indirect_noise.png}
\caption{\label{figure:noise}A screenshot showing noise in the indirect lighting.}
\end{figure}
\begin{figure}[htbp]\centering
\includegraphics[width=1.0\linewidth]{images/banding.png}
\caption{\label{figure:banding}A screenshot showing banding artefacts in the specular component.}{}
\end{figure}
